Fast, Scalable Phrase-Based SMT Decoding

### 2016-03-18 acl.2016/acl2016.tex
acl

\begin{abstract}
The utilization of statistical machine translation (SMT) has grown enormously over the last decade, many using open-source software developed by the NLP community. As commercial utilization has increased, there has been a pressing need that is optimized for their requirements. Specifically, faster phrase-based decoding, and more efficient utilization of modern multicore servers.

@@

SMT has been one of the outstanding success story from the NLP community in the last decade, progressing from a mostly research discipline to public useability via services such as Google Translate, Microsoft Translator Hub, as well as services and products built around offline products such as Language Weaver and the open-source Moses toolkit. The latter has spawned a cottage industry  encompassing a range of organizations and services from small language service providers seeking to reduce translation cost, to large inter-govermental organizations such as the EU and the UN that require high volume, high quality translation.

For high volume users, decoding is a largest and most critical part of the translation process which needs to be fast and efficient. However, it has been noticed that the Moses decoder, amongst others, is unable to efficiently use multiple CPU cores that are now common on modern servers~\cite{mfernandez2016boosting}. That is, the time taken to decode a test set does not substantial decrease when more cores are used, in fact, decoding time may increase when more cores are added. The problem will continue to grow as the commercial use of SMT increases and the number of CPU cores increases.

There have be speculation on the causes of the inefficiency as well as potential remedies. This paper is the first we know of that seeks to tackle this problem head on. We present an phrase-based decoder that is not only significantly faster than the Moses baseline for single-threaded operation, but is able to run multiple threads on multicore machines with only a slightly loss in linear speed. Model scores and functionality are compatible with Moses to aid comparison and ease of transition for users. All source code will be made available under an open-source license.

@@

There are a number of open-source SMT projects, most includes a decoder. The most well known is Moses~\cite{koehn-EtAl:2007:PosterDemo}, which supports phrase-based models, hierarchical phrase-based as well as various syntax-based models. Joshua~\cite{Joshua-Decoder} also supports hiearchical and syntax models and has recently supported phrase-based models. Phrasal supports a number of variants of the phrase-based model. CDEC~\cite{dyer2010cdec} supports hierarchical and syntactic models.

A number of the decoders support multithreading whilst others use alternative methods such as Hadoop or external scripts to parallelize decoding. We shall investigate the efficiency of using parallelizing decoding using the multi-processor approach. None of the decoder focus on multi-threads decoding.

~\newcite{mfernandez2016boosting} describes running multiple processes of the Moses decoder to increase speed, treating it as a black box within a parallelization framework.

Other prior work look to optimizing specific components of decoding. ~\newcite{Chiang:2007:cl} describes the cube-pruning and cube-growing algorithm for decoding which allows the tradeoff between speed and translation quality to the adjusted with a simgle parameter. ~\newcite{Heafield-kenlm} and ~\newcite{tanaka2013efficient} describes fast, efficient datastructures for language models. ~\newcite{zens2007efficient} describes an implementation of a phrase-table for an SMT decoder that is loaded on demand, reducing the initial loading time and memory requirements. ~\newcite{junczys_tsd_2012b} extends this by compressing the on-disk phrase table and lexicalized re-ordering model resulting in impressive speed gains over previous work.

~\newcite{Heafield-mtplz} is perhaps closest in intent to this work. This takes a holistic approach to decoding, describing a novel decoding algorithm which is focused on better decoding speed. It also describes a number of implementation details for faster decoding. However, the decoding algorithm is only able to incorporate one stateful feature function which precludes some of the useful decoding configurations which contains multiple stateful feature functions. It does not include a load-on-demand phrase table, therefore, cannot be used in a commercial environment where the phrase-table has not be filtered with a known test set for any realistic size phrase-table. Neither did this paper analyze the scalability of their work to multicore servers.

>>>>

\begin{abstract}
The utilization of statistical machine translation (SMT) has grown enormously over the last decade, many using open-source software developed by the NLP community. As commercial use has increased, there is need that for the software to be optimized for their requirements, in particular, faster phrase-based decoding and more efficient utilization of modern multicore servers.

@@

SMT has steadily progressed from a research discipline to public useability during the past decade via services such as Google Translate, Microsoft Translator Hub, as well as products and services for customized translation such as those from Language Weaver and the open-source Moses toolkit. The latter has spawned a cottage industry ranging from small language service providers seeking to reduce translation cost, to large inter-govermental agencies such as the EU and the UN that require high volume, in-house machine translation.

For high volume users, decoding is a most time critical part of the translation process. A common strategy to increase decoding throughput is to make use of the multiple cores that are now ubiquitous in todays servers. However, it has been noticed that the Moses decoder, amongst others, is unable to efficiently use multiple~\cite{mfernandez2016boosting}. That is, decoding speed does not substantial increase when more cores are used, in fact, it may actually \emph{decrease} when more cores are added.

There have be speculation on the causes of the inefficiency as well as potential remedies. This paper is the first we know of that seeks to tackle this problem head on. We present an phrase-based decoder that is not only significantly faster than the Moses baseline for single-threaded operation, but is able to run multiple threads on multicore machines with only a slightly loss in linear speed. Model scores and functionality are compatible with Moses to aid comparison and ease transition for users. All source code will be made available under an open-source license.

@@

There are a number of open-source SMT projects, most includes a decoder. The most well known is Moses~\cite{koehn-EtAl:2007:PosterDemo}, which supports phrase-based models, hierarchical phrase-based as well as various syntax-based models. Joshua~\cite{Joshua-Decoder} also supports hierarchical and syntax models and has recently supported phrase-based models. Phrasal supports a number of variants of the phrase-based model. CDEC~\cite{dyer2010cdec} supports hierarchical and syntactic models.

A number of the decoders support multithreading whilst others use alternative methods such as Hadoop or external scripts to parallelize decoding.

Most prior work on increasing decoding speed look to optimizing specific components of decoding. ~\newcite{Chiang:2007:cl} describes the cube-pruning and cube-growing algorithm for decoding which allows the tradeoff between speed and translation quality to the adjusted with a simgle parameter. ~\newcite{Heafield-kenlm} and ~\newcite{tanaka2013efficient} describes fast, efficient datastructures for language models. ~\newcite{zens2007efficient} describes an implementation of a phrase-table for an SMT decoder that is loaded on demand, reducing the initial loading time and memory requirements. ~\newcite{junczys_tsd_2012b} extends this by compressing the on-disk phrase table and lexicalized re-ordering model resulting in impressive speed gains over previous work. ~\newcite{mfernandez2016boosting} describes running multiple processes of the Moses decoder to increase speed, treating it as a black box within a parallelization framework.

~\newcite{Heafield-mtplz} is perhaps closest in intent to our work. This takes a holistic approach to decoding, describing a novel decoding algorithm which is focused on better decoding speed. However, the decoding algorithm is only able to incorporate one stateful feature function which precludes some of the useful decoding configurations which contains multiple stateful feature functions. It does not include a load-on-demand phrase table, therefore, cannot be used in a commercial environment with any realistic size models. Neither did this paper analyze the scalability of their work to multicore servers.
